---
title: "Analysis and Visualization of Data for Bach Visualization Project"
output:
  html_document:
    df_print: paged
---

This notebook conducts analysis of the questionnaire presented to the live and livestreaming audiences of MusicLab Copenhagen and a follow-up sample of online participants who responded to continuous rating scales and the same survey that was provided to the concert audience. The survey can be found here: https://oslopsych.az1.qualtrics.com/jfe/form/SV_9n5uwhPGOHJqqHQ

```{r}
citation()
citation("psych")
citation("lme4")
```


# Import libraries

```{r libraries}
packages = c("psych", "readxl","readr", "magrittr", "reshape2","ggpubr", "ltm","janitor", "PerformanceAnalytics", "tidyverse", "lme4", "ggforce", "cowplot", "car") # car: companionto applied regression

## Now load or install&load all
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)

source("useful_functions.R")
```

# Load Data
## Live Concert
```{r}
df.full<-readRDS(file = "../output/Prepared_Data.Rda")

# rename ParticipantCode to Pt_ID because that is the name I began by using here.
names(df.full)[1]<-"Pt_ID"
```

### save the df.full into an Rda for use with the 

```{r}
dat<-df.full%>%
  select(Pt_ID, group, age, EC,AIMS,musician_status, fan, perceive_Bach:style_Bach)%>%
  filter(group == "Live")%>%
  select(-group)
```

## Follow up study
```{r}
# raw
rawurlfile = "https://raw.githubusercontent.com/finn42/Bach_Viz_Responses/main/data_input/raw.csv"
raw<-read_csv(url(rawurlfile))

# edited
editedurlfile="https://raw.githubusercontent.com/finn42/Bach_Viz_Responses/main/data_input/edited.csv"
edited<-read_csv(url(editedurlfile))
```

### Save the edited dataframe for use in an anonymized code example
```{r}
#save(edited, file = "Online_Study_Data.rda")
```

# Organize data
## Online follow-up

Add ID column
```{r}
edited<-tibble::rowid_to_column(edited, "Pt_ID")
```

**Sort data into simple and complex**
Half of the participants viewed the simple first and half the complex first. You can determine who received which stimulus based on if they have NA in complex1 and not NA in simple 1 then they watched the simple video first. 

- I will need to also keep track of if they viewed the simple or the complex first

Label as simple or complex first or second. 

Note: participants could have seen any combination of complex and simple.(e.g. 1C, 1S or 2C or 2S)
```{r}
edited<-edited%>%mutate(type1 = case_when(
    !is.na(simple1) ~ "Simple",
    !is.na(complex1) ~ "Complex",
    ))

edited<-edited%>%mutate(type2 = case_when(
    !is.na(simple2) ~ "Simple",
    !is.na(complex2) ~ "Complex",
    ))
```

Drop the unnecessary data of continuous ratings
```{r}
edited<-edited%>%select(!c(complex1, simple1, complex2, simple2))
```

Desired colnames
```{r}
col_names<-names(dat)[7:length(dat)]
print(col_names)
```
Which cols correspond to the first video and which the second?
```{r}
# questions on the first video 
transpose(raw[1, c('Q3', 'Q4', 'Q5', 'Q6', 'Q7', 'Q8', 'Q9','Q11', 'Q12')]) # questions corresponding to column labels

# questions on the second video 
transpose(raw[1, c('Q21', 'Q22', 'Q23', 'Q24', 'Q25', 'Q26', 'Q27')]) # questions corresponding to column labels

# questions on personal chars 
transpose(raw[1, c('Q49', 'Q52', 'Q53', 'Q54', 'Q58')]) # questions corresponding to column labels
```

Rename personal Characteristic cols
```{r}
colnames(edited)[which(names(edited) == "Q49")] <- "quiz"
colnames(edited)[which(names(edited) == "Q52")] <- "age"
colnames(edited)[which(names(edited) == "Q53")] <- "gender"
colnames(edited)[which(names(edited) == "Q54")] <- "musician_status"
```

We also had a question that was a quiz question to truly test participants' Grasping of the visualization:

Please select the statement that is most correct.
X The color of the discs corresponded to the emotion in the music.
X Each line corresponded to a different string on the violin.
--> Each disc corresponded to a different instrument.
X The color of the lines corresponded to the harmonic color of the music.

Calculate quiz result
```{r}
edited$quiz_result<-0
edited$quiz_result[edited$quiz==3]<-1
```

How many participants answered the quiz correctly?
```{r}
edited%>%group_by(quiz)%>%summarise(n())
edited%>%group_by(quiz_result)%>%summarise(n())
```


Rename the columns based on if they viewed the simple or complex first

To do this, I need to create new dataframes.
```{r}
# first set of Qs
first_video<-edited%>%select(Pt_ID,type1, Q3:Q12)
# second set of Qs
second_video<-edited%>%select(Pt_ID, type2, Q21:Q27)
```

Rename cols
```{r}
names(first_video)[3:11]<-col_names
names(second_video)[3:9]<-col_names[1:(length(col_names)-2)]
```

Pivot longer
```{r}
first_long<-first_video%>%
  pivot_longer(!c(Pt_ID,type1), names_to = "question", values_to = "response")

second_long<-second_video%>%
  pivot_longer(!c(Pt_ID,type2), names_to = "question", values_to = "response")
```

Rename type
```{r}
colnames(first_long)[2]<-"type"
colnames(second_long)[2]<-"type"

first_long$order<-"first"
second_long$order<-"second"

long<-rbind(first_long, second_long)
long$group<-"Online"
```

Pivot wider
```{r}
pc<-edited%>%select(Pt_ID, age, musician_status)
online_df<-full_join(long, pc, by = "Pt_ID")
online_df.wide<-online_df%>%pivot_wider(names_from = question, values_from= response)
```

Factor musician status
```{r}
online_df.wide$musician_status<-factor(online_df.wide$musician_status)
online_df.wide$musician_status<-fct_collapse(online_df.wide$musician_status,"Tone-deaf" = "1", "Nonmusician" = "2", "Music-loving nonmusician" ="3", "Amateur musician" = "4", "Serious amateur musician" = "5", "Semiprofessional musician"= "6", "Professional musician" = "7")

online_df.wide$Pt_ID<-factor(online_df.wide$Pt_ID)
```

The scale was 1-7 for the familiar and style items therefore subtract 1 to be on the same 0-6 scale as in the live data
```{r}
# if statement to make sure this does not happen twice. 
if(max(online_df.wide$familiar_Bach, na.rm = TRUE) == 7){
  online_df.wide$familiar_Bach<-online_df.wide$familiar_Bach-1
}
if(max(online_df.wide$style_Bach, na.rm = TRUE) == 7){
  online_df.wide$style_Bach<-online_df.wide$style_Bach-1
}
```

## Live
Remove the cols that are not in the online data
```{r}
names(dat)
live_dat<-dat%>%select(-EC, -AIMS, -fan)

live_dat$group<-"Live"
live_dat$type<-"Complex"
live_dat$order<-"live"
```
## Join together
Join the online and the concert results. 
```{r}
full_data<-full_join(live_dat, online_df.wide)
```
Convert to long again
```{r}
other_vars<-c("Pt_ID", "age", "musician_status", "group", "type","order")

full_data.long<-full_data%>%
  pivot_longer(!other_vars, names_to = "question", values_to = "response")

full_data.long$question<-factor(full_data.long$question, levels = c("perceive_Bach", "instruments_Bach","understanding_Bach", "augment_Bach", "disturb_Bach", "enhance_Bach", "emotionally_moved_Bach", "familiar_Bach","style_Bach"), labels = c("Perceive", "Instruments","Understanding", "Augment", "Disturb", "Enhance", "Moved", "Familiar","Style"))

scale_five<-c("Perceive", "Instruments","Understanding", "Augment", "Disturb", "Enhance", "Moved")
full_data.long$facet<-"7-item scale"
full_data.long$facet[full_data.long$question %in% scale_five]<-"5-item scale"

full_data.long$facet_group[full_data.long$type == "Simple"]<-"Online - Simple"
full_data.long$facet_group[full_data.long$type == "Complex"]<-"Online - Complex"
full_data.long$facet_group[full_data.long$group == "Live"]<-"Live"

full_data.long$facet_group<-factor(full_data.long$facet_group, levels = c("Live", "Online - Simple", "Online - Complex"))

# Why are there NA in facet_group?
## Because Participant 32 in the online experiment did not respond in the live rating of the second piece at all therefore we can remove the NAs that exist for this participant
full_data.long<-full_data.long%>%filter(!(Pt_ID == "32" & is.na(type)))
```
# Visualize
### Complex: Live versus Online - Stacked bar chart
```{r}
Title = "Simple Bach Visualization"

plot_data<-full_data.long%>%filter(question %in% scale_five)%>%filter(type == "Complex")%>%filter(!is.na(response))

plot_data$Response<-factor(plot_data$response, levels = c("5", "4", "3", "2", "1"), labels = c("5 Totally", "4", "3", "2", "1 Not at all"))

# Put the sample size in the subtitle
n = plot_data%>%group_by(group)%>%summarise(n = n_distinct(Pt_ID))

Subtitle = paste0("Live (n = ", n$n[n$group == "Live"], ") and Online (n = ", n$n[n$group == "Online"],")")

plot_data%>%ggplot(aes(x = question, fill = Response))+
  geom_bar(position = "fill")+
  labs(title=Title, subtitle = Subtitle)+
  scale_fill_brewer(palette = "Reds", direction = -1) +
  theme_minimal()+
  xlab("Question")+
  ylab("Proportion")+
  facet_grid(rows = vars(group)) +
  theme(axis.text.x = element_text(angle = 30)) #, vjust = 0.5, hjust=1))

graphname<-paste0("../plots/Bar_",Title,"subtitle.png")

ggsave(graphname, 
       width = 15, 
       height = 10, 
       units = 'cm', 
       dpi = 500)
  
```


## Online: Complex versus simple

```{r}
Title = "Online Bach Visualization"

plot_data<-full_data.long%>%filter(question %in% scale_five)%>%filter(group == "Online")

plot_data$Response<-factor(plot_data$response, levels = c("5", "4", "3", "2", "1"), labels = c("5 Totally", "4", "3", "2", "1 Not at all"))

plot_data$type<-factor(plot_data$type, levels = c("Simple", "Complex"))

# Put the sample size in the subtitle
n = plot_data%>%group_by(type)%>%summarise(n = n_distinct(Pt_ID))

Subtitle = paste0("Simple (n = ", n$n[n$type == "Simple"], ") and Complex (n = ", n$n[n$type == "Complex"],")")

plot_data%>%ggplot(aes(x = question, fill = Response))+
  geom_bar(position = "fill")+
  labs(title=Title, subtitle = Subtitle)+
  scale_fill_brewer(palette = "Reds", direction = -1) +
  theme_minimal()+
  xlab("Question")+
  ylab("Proportion")+
  facet_grid(rows = vars(type)) +
  theme(axis.text.x = element_text(angle = 30)) #, vjust = 0.5, hjust=1))

graphname<-paste0("../plots/Bar_Complex_Simple",Title,".png")

ggsave(graphname, 
       width = 15, 
       height = 10, 
       units = 'cm', 
       dpi = 500)
```

## Single chart: Live, Simple, Complex
When comparing live to simple and simple to complex, we are having a repetition of information. 
Additionally, we could make separate plots to give them separate colors and then combine the plots with cowplot for example. 

```{r}
Title = "Bach Visualization"

# Live
plot_data<-full_data.long%>%filter(question %in% scale_five)%>%filter(group == "Live")%>%filter(!is.na(response))

n_Live = plot_data%>%summarise(n = n_distinct(Pt_ID))

Subtitle = paste0("Live Complex (n = ", n_Live$n, ")")

plot_data$Response<-factor(plot_data$response, levels = c("5", "4", "3", "2", "1"), labels = c("5 Totally", "4", "3", "2", "1 Not at all"))

p_live<-plot_data%>%ggplot(aes(x = question, fill = Response))+
  geom_bar(position = "fill")+
  labs(title=Title, subtitle = Subtitle)+
  scale_fill_brewer(palette = "Reds", direction = -1, name = "") +
  theme_minimal()+
  xlab("")+
  ylab("Proportion")+
  theme(axis.text.x = element_blank())
  #theme(axis.text.x = element_text(angle = 30)) #, vjust = 0.5, hjust=1))

# Complex
plot_data<-full_data.long%>%filter(question %in% scale_five)%>%filter(group == "Online" & type == "Complex")%>%filter(!is.na(response))

n_Complex = plot_data%>%summarise(n = n_distinct(Pt_ID))

Subtitle = paste0("Online Complex (n = ", n_Complex$n, ")")

plot_data$Response<-factor(plot_data$response, levels = c("5", "4", "3", "2", "1"), labels = c("5 Totally", "4", "3", "2", "1 Not at all"))

p_complex<-plot_data%>%ggplot(aes(x = question, fill = Response))+
  geom_bar(position = "fill")+
  labs(subtitle = Subtitle)+
  scale_fill_brewer(palette = "YlGn", direction = -1, name = "") + # same palette to reflect that they are the same group!@
  theme_minimal()+
  xlab("")+
  ylab("Proportion")+
  theme(axis.text.x = element_blank())
  #theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1))

p_complex

# Simple
plot_data<-full_data.long%>%filter(question %in% scale_five)%>%filter(group == "Online" & type == "Simple")%>%filter(!is.na(response))

n_Simple = plot_data%>%summarise(n = n_distinct(Pt_ID))

Subtitle = paste0("Online Simple (n = ", n_Simple$n,")")

plot_data$Response<-factor(plot_data$response, levels = c("5", "4", "3", "2", "1"), labels = c("5 Totally", "4", "3", "2", "1 Not at all"))

p_simple<-plot_data%>%ggplot(aes(x = question, fill = Response))+
  geom_bar(position = "fill")+
  labs(subtitle = Subtitle)+
  scale_fill_brewer(palette = "YlGn", direction = -1, name = "") +
  theme_minimal()+
  xlab("")+
  ylab("Proportion")+
  #theme(axis.text.x = element_blank())
  theme(axis.text.x = element_text(angle = 30)) #, vjust = 0.5, hjust=1))

plot_grid(
  p_live, p_complex, p_simple,
  labels = "AUTO", ncol = 1,
  rel_heights = c(1.1,1,1.2)
)

graphname<-paste0("../plots/Bar_3_",Title,".png")

ggsave(graphname, 
       width = 15, 
       height = 20, 
       units = 'cm', 
       dpi = 500)
```

## Single chart: Live, First: Simple, Complex; Second: Simple, Complex
The order of presentation may be important therefore make a plot that splits information into simple first, simple second, complex first, complex second

```{r}
Title = "Bach Visualization"

# Live
plot_data<-full_data.long%>%filter(question %in% scale_five)%>%filter(group == "Live")%>%filter(!is.na(response))

n_Live = plot_data%>%summarise(n = n_distinct(Pt_ID))

Subtitle = paste0("Live: Complex (n = ", n_Live$n, ")")

plot_data$Response<-factor(plot_data$response, levels = c("5", "4", "3", "2", "1"), labels = c("5 Totally", "4", "3", "2", "1 Not at all"))

p_live<-plot_data%>%ggplot(aes(x = question, fill = Response))+
  geom_bar(position = "fill")+
  labs(title=Title, subtitle = Subtitle)+
  scale_fill_brewer(palette = "Reds", direction = -1, name = "") +
  theme_minimal()+
  xlab("")+
  ylab("Proportion")+
  theme(axis.text.x = element_blank())
  #theme(axis.text.x = element_text(angle = 30)) #, vjust = 0.5, hjust=1))

# Complex First
plot_data<-full_data.long%>%filter(question %in% scale_five)%>%filter(group == "Online" & type == "Complex" & order == "first")%>%filter(!is.na(response))

n_Complex_First = plot_data%>%summarise(n = n_distinct(Pt_ID))

Subtitle = paste0("Online: Complex First (n = ", n_Complex_First$n, ")")

plot_data$Response<-factor(plot_data$response, levels = c("5", "4", "3", "2", "1"), labels = c("5 Totally", "4", "3", "2", "1 Not at all"))

p_complex_first<-plot_data%>%ggplot(aes(x = question, fill = Response))+
  geom_bar(position = "fill")+
  labs(subtitle = Subtitle)+
  scale_fill_brewer(palette = "YlGn", direction = -1, name = "") + # same palette to reflect that they are the same group!@
  theme_minimal()+
  xlab("")+
  ylab("Proportion")+
  theme(axis.text.x = element_blank())
  #theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1))

p_complex_first

# Simple First
plot_data<-full_data.long%>%filter(question %in% scale_five)%>%filter(group == "Online" & type == "Simple" & order == "first")%>%filter(!is.na(response))

n_Simple_first = plot_data%>%summarise(n = n_distinct(Pt_ID))

Subtitle = paste0("Online: Simple First (n = ", n_Simple_first$n,")")

plot_data$Response<-factor(plot_data$response, levels = c("5", "4", "3", "2", "1"), labels = c("5 Totally", "4", "3", "2", "1 Not at all"))

p_simple_first<-plot_data%>%ggplot(aes(x = question, fill = Response))+
  geom_bar(position = "fill")+
  labs(subtitle = Subtitle)+
  scale_fill_brewer(palette = "YlGn", direction = -1, name = "") +
  theme_minimal()+
  xlab("")+
  ylab("Proportion")+
  theme(axis.text.x = element_blank())
  #theme(axis.text.x = element_text(angle = 30)) #, vjust = 0.5, hjust=1))

# Complex Second
plot_data<-full_data.long%>%filter(question %in% scale_five)%>%filter(group == "Online" & type == "Complex" & order == "second")%>%filter(!is.na(response))

n_Complex_Second = plot_data%>%summarise(n = n_distinct(Pt_ID))

Subtitle = paste0("Online: Complex Second (n = ", n_Complex_Second$n, ")")

plot_data$Response<-factor(plot_data$response, levels = c("5", "4", "3", "2", "1"), labels = c("5 Totally", "4", "3", "2", "1 Not at all"))

p_complex_second<-plot_data%>%ggplot(aes(x = question, fill = Response))+
  geom_bar(position = "fill")+
  labs(subtitle = Subtitle)+
  scale_fill_brewer(palette = "YlGn", direction = -1, name = "") + # same palette to reflect that they are the same group!@
  theme_minimal()+
  xlab("")+
  ylab("Proportion")+
  theme(axis.text.x = element_blank())
  #theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1))

p_complex_second

# Simple First
plot_data<-full_data.long%>%filter(question %in% scale_five)%>%filter(group == "Online" & type == "Simple" & order == "second")%>%filter(!is.na(response))

n_Simple_second = plot_data%>%summarise(n = n_distinct(Pt_ID))

Subtitle = paste0("Online: Simple Second (n = ", n_Simple_second$n,")")

plot_data$Response<-factor(plot_data$response, levels = c("5", "4", "3", "2", "1"), labels = c("5 Totally", "4", "3", "2", "1 Not at all"))

p_simple_second<-plot_data%>%ggplot(aes(x = question, fill = Response))+
  geom_bar(position = "fill")+
  labs(subtitle = Subtitle)+
  scale_fill_brewer(palette = "YlGn", direction = -1, name = "") +
  theme_minimal()+
  xlab("")+
  ylab("Proportion")+
  #theme(axis.text.x = element_blank())
  theme(axis.text.x = element_text(angle = 30)) #, vjust = 0.5, hjust=1))

plot_grid(
  p_live, p_complex_first, p_simple_first, p_complex_second, p_simple_second,
  labels = "AUTO", ncol = 1,
  rel_heights = c(1.15,1,1,1,1.3)
)

graphname<-paste0("../plots/Bar_5_",Title,".png")

ggsave(graphname, 
       width = 15, 
       height = 25, 
       units = 'cm', 
       dpi = 500)
```


## Online: SS, SC, CS, CC
Add subgroup class: SS, SC, CS, CC
```{r}
necessary_info<-online_df.wide%>%select(Pt_ID, order, type)
necessary_info.wide<-necessary_info%>%pivot_wider(names_from = order, values_from = type)
necessary_info.wide$Subgroup[necessary_info.wide$first == "Simple"&necessary_info.wide$second == "Simple"]<-"SS"
necessary_info.wide$Subgroup[necessary_info.wide$first == "Simple"&necessary_info.wide$second == "Complex"]<-"SC"
necessary_info.wide$Subgroup[necessary_info.wide$first == "Complex"&necessary_info.wide$second == "Simple"]<-"CS"
necessary_info.wide$Subgroup[necessary_info.wide$first == "Complex"&necessary_info.wide$second == "Complex"]<-"CC"

subgroups<-necessary_info.wide%>%select(Pt_ID, Subgroup)

#join with full_data.long
full_data.long<-full_data.long%>%full_join(subgroups, by = "Pt_ID")
```

```{r}
Title = "Online Bach Visualization"

plot_data<-full_data.long%>%filter(question %in% scale_five)%>%filter(group == "Online")
plot_data$Subgroup<-factor(plot_data$Subgroup, levels = c("SS", "SC", "CS", "CC"))

# Put the sample size in the subtitle

n = plot_data%>%group_by(Subgroup)%>%summarise(n = n_distinct(Pt_ID))

Subtitle = paste0("SS =", n$n[n$Subgroup == "SS"], ", SC = ", n$n[n$Subgroup == "SC"],", CS = ", n$n[n$Subgroup == "CS"],", CC =", n$n[n$Subgroup =="CC"])

summary_data<-plot_data%>%group_by(Subgroup, order, question)%>%summarise(Response = mean(response))%>%filter(!is.na(Subgroup))

summary_data%>%ggplot(aes(x = order, y = Response, color = Subgroup, group = Subgroup))+
  geom_line()+
  #geom_point()+
  labs(title=Title, subtitle = Subtitle)+
  theme_DSQ()+
  xlab("Question")+
  ylab("Response")+
  facet_wrap(. ~ question) +
  theme(axis.text.x = element_text(angle = 30)) #, vjust = 0.5, hjust=1))

graphname<-paste0("../plots/Line_",Title,".png")

ggsave(graphname, 
       width = 15, 
       height = 15, 
       units = 'cm', 
       dpi = 500)
```
## Familiar with Piece and Style
Visualize differences between the audiences on their familiarity
Were you familiar with the music that you heard? 0 = I have never heard it before - 6 - I am very familiar with the music
Were you familiar with this style of music (Bach fugues)? 0 = I have never heard it before  - 6 = I am very familiar with this type of music
```{r}
plot_data<-full_data.long%>%filter(!question %in% scale_five)%>%filter(! is.na(response))

plot_data$Response<-factor(plot_data$response, levels = c("6", "5", "4", "3", "2", "1", "0"), labels = c("6","5", "4", "3", "2", "1", "0"))

Title= "Familiar with Piece and Style"
Subtitle = "Live"

fam_live<-plot_data%>%filter(group == "Live")%>%ggplot(aes(x = question, fill = Response))+
  geom_bar(position = "fill")+
  labs(title = Title, subtitle = Subtitle)+
  scale_fill_brewer(palette = "Reds", direction = -1, name = "") + # same palette to reflect that they are the same group!@
  theme_minimal()+
  #facet_grid(rows = vars(group)) +
  xlab("")+
  ylab("Proportion")+
  theme(axis.text.x = element_blank())
  #theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1))

Subtitle = "Online"
fam_online<-plot_data%>%filter(group == "Online")%>%ggplot(aes(x = question, fill = Response))+
  geom_bar(position = "fill")+
  labs(subtitle = Subtitle)+
  scale_fill_brewer(palette = "YlGn", direction = -1, name = "") + # same palette to reflect that they are the same group!@
  theme_minimal()+
  #facet_grid(rows = vars(group)) +
  xlab("")+
  ylab("Proportion")+
  #theme(axis.text.x = element_blank())
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1))

plot_grid(
  fam_live, fam_online,
  labels = "AUTO", ncol = 1
  #rel_heights = c(1.1,1,1.2)
)

graphname<-paste0("../plots/Bar_",Title,".png")

ggsave(graphname, 
       width = 10, 
       height = 15, 
       units = 'cm', 
       dpi = 500)
```


## Violin Plot: Live only
```{r}
# transform the response scale of the familiarity questions to 1-5 
## 0-->1(*5/7 + 1)
#dat<-dat%>%mutate(style_Bach_t = (style_Bach+1)*5/7,familiar_Bach_t = (familiar_Bach+1)*5/7)

# Or just make a facet grid so that you can have different y scales? See below
#facet<-c(1,1,1,1,1,1,1,2,2)

personal_chars<-c("Pt_ID", "age", "EC", "AIMS","musician_status", "fan")

dat_long<-dat%>%
  pivot_longer(!all_of(personal_chars), names_to = "question", values_to = "response")

dat_long$question<-factor(dat_long$question, levels = c("perceive_Bach", "instruments_Bach","understanding_Bach", "augment_Bach", "disturb_Bach", "enhance_Bach", "emotionally_moved_Bach", "familiar_Bach","style_Bach"), labels = c("Perceive", "Instruments","Understanding", "Augment", "Disturb", "Enhance", "Moved", "Familiar","Style"))

scale_five<-c("Perceive", "Instruments","Understanding", "Augment", "Disturb", "Enhance", "Moved")
dat_long$facet<-"7-item scale"
dat_long$facet[dat_long$question %in% scale_five]<-"5-item scale"
```

### Facet: 5 and 7-item scale 
```{r}
Title = "Bach Visualization"

dat_long%>%ggplot(aes(x = question, y = response))+
  geom_violin(show.legend= FALSE, fill = "#b34036")+
  labs(title=Title)+
  theme_DSQ()+
  xlab("Question")+
  ylab("Response")+
  #facet_wrap(~facet, scales = "free")+
  #facet_grid(. ~ facet, scales = "free", space='free') +
  ggforce::facet_row(vars(facet), scales = 'free', space = 'free')+
  theme(axis.text.x = element_text(angle = 30)) #, vjust = 0.5, hjust=1))

graphname<-paste0("../plots/Violin_",Title, ".png")
ggsave(graphname, 
       width = 15, 
       height = 9, 
       units = 'cm', 
       dpi = 500)
```

# Statistical Analysis
## Participants

How many participants participated?

86 participants answered the Bach questionnaire completely. 
1 participant was missing one response (perceive), 
```{r}
# ID pts who did not answer any Bach survey questions

# This line of code drops any participant with even 1 NA in the Bach questionnaire. This indicates that 86 participants answered the Bach questionnaire completely. 
#demo_dat<-full_data%>%drop_na(contains("Bach"))%>%select(Pt_ID, group, age, musician_status)%>%distinct()

# Instead we need to drop rows with all NA in the cols
count_dat<-full_data%>%select(Pt_ID, group, contains("Bach"))
count_dat<-count_dat[rowSums(is.na(count_dat)) != ncol(count_dat)-2,] # -2 because need to retain Pt_ID and group in the df

count_dat%>%group_by(group)%>%summarise(count = n())

# check which vars the 3 participants missed.
live_count_dat<-count_dat%>%filter(group == "Live")
## 1 missed filling "perceive", 2 missed filling "disturb". These participants had no other NAs.
```

Compare demographics between audiences.
Compare age and musician status between live and online groups.
```{r}
demo_dat<-full_data%>%drop_na(contains("Bach"))%>%select(Pt_ID, group, age, musician_status)%>%distinct()

#age
demo_dat%>%group_by(group)%>%summarise(mean = mean(age, na.rm = TRUE), sd = sd(age, na.rm = TRUE))

#musician status
test<-demo_dat%>%group_by(group,musician_status)%>%summarise(count = n())
test2<-pivot_wider(test, names_from = group, values_from = count )
write.table(test2, file = "musician_status_Bach_live_v_online.txt", sep = ",", quote = FALSE, row.names = F)

library(xtable)
print(xtable(test2, type = "latex"), file = "musician_status_Bach_live_v_online.tex")
```

### Age: Two-sample t-test
The live audience was significantly older than the online audience.
```{r}
t.test(demo_dat$age~demo_dat$group)
```

### Musician status (ordinal) Kruskal Wallis test
The Kruskal Wallis test is used when you have one independent variable with two or more levels and an ordinal dependent variable.

The live audience was less musically sophisticated than the online participants
```{r}
ggplot(demo_dat, aes(x = musician_status))+geom_bar()+facet_grid(cols = vars(group)) # these distributions don't really look the same 

kruskal.test(musician_status~group, data = demo_dat)
```

## Correlations
### Live 
take a glance with the performance analytics correlation chart

Reload data
```{r}
personal_chars<-c("Pt_ID", "age", "EC", "AIMS","musician_status", "fan")

dat<-df.full%>%
  select(Pt_ID, group, age, EC,AIMS,musician_status, fan, perceive_Bach:style_Bach)%>%
  filter(group == "Live")%>%
  select(-group)
```

```{r}
cor_data_Bach<-dat%>%select(!Pt_ID)
```

```{r}
#change musician_status to ranking
cor_data_Bach$musician_status<-as.numeric(cor_data_Bach$musician_status)

#Cross correlation
chart.Correlation(cor_data_Bach, method = "pearson")
chart.Correlation(cor_data_Bach, method = "spearman")
chart.Correlation(cor_data_Bach, method = "kendall")
```

```{r}
#select only complete
cor_data_Bach_nona<-na.omit(cor_data_Bach) # however maybe you could retain more data if you simply got rid of the age col since this seems to be what is missing most often.
names(cor_data_Bach_nona)<-c("Age", "Empathic Concern","AIMS", "Musician Status", "Fan-status", "Perceive", "Instruments","Understanding", "Augment", "Disturb", "Enhance", "Moved", "Familiar","Style")

# check number of samples
samples<-nrow(cor_data_Bach_nona)
colours_DSQ<-c("#2d769a", "white","#b34036")
title = ""
#title = "Bach Visualization"
subtitle = paste0("Kendall Correlations (BH adj) (n = ", samples, ")")
chart.correlation(cor_data_Bach_nona, colours_DSQ, title, subtitle) #moved the cross-correlation code to the useful_functions script to make this more tidy.

graphname<-paste0("../plots/Kendall-Correlations_BachVis.png")
ggsave(graphname, 
       width = 10, 
       height = 9, 
       units = 'cm', 
       dpi = 500)
```
### Remove age
from the Corr chart shwoign live participants to be able to retain more participants
```{r}
cor_data_Bach<-dat%>%select(!age)

#select only complete
cor_data_Bach_nona<-na.omit(cor_data_Bach) # however maybe you could retain more data if you simply got rid of the age col since this seems to be what is missing most often.
names(cor_data_Bach_nona)<-c("Age", "Empathic Concern","AIMS", "Musician Status", "Fan-status", "Perceive", "Instruments","Understanding", "Augment", "Disturb", "Enhance", "Moved", "Familiar","Style")

# check number of samples
samples<-nrow(cor_data_Bach_nona)
colours_DSQ<-c("#2d769a", "white","#b34036")
title = ""
#title = "Bach Visualization"
subtitle = paste0("Kendall Correlations (BH adj) (n = ", samples, ")")
chart.correlation(cor_data_Bach_nona, colours_DSQ, title, subtitle) #moved the cross-correlation code to the useful_functions script to make this more tidy.

graphname<-paste0("../plots/Kendall-Correlations_noAge_BachVis.png")
ggsave(graphname, 
       width = 10, 
       height = 9, 
       units = 'cm', 
       dpi = 500)
```
### Live and Online

```{r}
cor_data_Bach<-full_data%>%select(age, musician_status, perceive_Bach:style_Bach) # note not Pt_ID, group, 

#change musician_status to ranking
cor_data_Bach$musician_status<-as.numeric(cor_data_Bach$musician_status)

#select only complete
cor_data_Bach_nona<-na.omit(cor_data_Bach) # however maybe you could retain more data if you simply got rid of the age col since this seems to be what is missing most often.
names(cor_data_Bach_nona)<-c("Age", "Musician Status", "Perceive", "Instruments","Understanding", "Augment", "Disturb", "Enhance", "Moved", "Familiar","Style")

# check number of samples
samples<-nrow(cor_data_Bach_nona)
colours_DSQ<-c("#2d769a", "white","#b34036")
title = ""
#title = "Bach Visualization"
subtitle = paste0("Kendall Correlations (BH adj) (n = ", samples, ")")
chart.correlation(cor_data_Bach_nona, colours_DSQ, title, subtitle) #moved the cross-correlation code to the useful_functions script to make this more tidy.

graphname<-paste0("../plots/Kendall-Correlations-Live_and_Online_BachVis.png")
ggsave(graphname, 
       width = 10, 
       height = 9, 
       units = 'cm', 
       dpi = 500)

```

### Online (Controlled Experiment) Only
Without the age variable

Musical sophistication: Zhang and Schubert 2019
Perceive: 
```{r}
cor_data_Bach<-full_data%>%filter(group == "Online")%>%select(musician_status, perceive_Bach:style_Bach) # note not Pt_ID, group, #age, 

#change musician_status to ranking
cor_data_Bach$musician_status<-as.numeric(cor_data_Bach$musician_status)

#select only complete
cor_data_Bach_nona<-na.omit(cor_data_Bach) # however maybe you could retain more data if you simply got rid of the age col since this seems to be what is missing most often.
names(cor_data_Bach_nona)<-c("Musical Sophistication", "Perceive", "Instruments","Understanding", "Augment", "Disturb", "Enhance", "Moved", "Familiar","Style") # "Age", 

# check number of samples
samples<-nrow(cor_data_Bach_nona)
colours_DSQ<-c("#2d769a", "white","#b34036")
title = ""
#title = "Bach Visualization"
subtitle = paste0("Kendall Correlations (BH adj) (n = ", samples, ")")
chart.correlation(cor_data_Bach_nona, colours_DSQ, title, subtitle) #moved the cross-correlation code to the useful_functions script to make this more tidy.

graphname<-paste0("../plots/Kendall-Correlations-Online_BachVis.png")
ggsave(graphname, 
       width = 10, 
       height = 9, 
       units = 'cm', 
       dpi = 500)

```


## Reliability: Omega
### Live
```{r}
visualization_items_live<-dat%>%select(-personal_chars, -familiar_Bach, -style_Bach)
omega(visualization_items_live)
```
### Online only
```{r}
visualization_items_online<-full_data%>%filter(group == "Online")%>%select(perceive_Bach:emotionally_moved_Bach)
omega(visualization_items_online)
```

### Live and Online
```{r}
visualization_items_all<-full_data%>%select(perceive_Bach:emotionally_moved_Bach)
omega(visualization_items_all)
```

## Factor analysis
To reduce dimensionality of the questions on the visualization so that we can then do model testing to understand the impact of musical expertise and familiarity on the experience.

### Live and Online
Redo the FA with full_data: live and online

```{r}
data_for_fa<-full_data%>%select(Pt_ID, perceive_Bach:emotionally_moved_Bach)
fa_dat<-data_for_fa%>%select(perceive_Bach:emotionally_moved_Bach)
```

```{r}
pairs.panels(fa_dat)
# Test for the number of factors using parallel analysis or very simple structure
fa.parallel(fa_dat) #this suggests that there are 2 factors and 1 component
vss(fa_dat) # this suggests that the complexity 1 solution achieves max at 1 factor and complexity 2 solution achieve maximum at 2 factors
```
#### One-factor Solution
```{r}
visualization_factor_analysis_1<-fa(fa_dat, 1)
visualization_factor_analysis_1
visualization_factor_analysis_1$Vaccounted
plot(visualization_factor_analysis_1)
fa.diagram(visualization_factor_analysis_1)
```
#### Two-factor Solution
It is interesting to see its structure because now there is a factor for understanding and an emotional/enhancing factor. But the understanding and enhancing factors are highly correlated.
```{r}
visualization_factor_analysis_2<-fa(fa_dat, 2)
visualization_factor_analysis_2
visualization_factor_analysis_2$Vaccounted
plot(visualization_factor_analysis_2)
fa.diagram(visualization_factor_analysis_2)
```

March 3rd, 2025: Compute metrics for assessing whether the dataset is appropriate for factor analysis as recommended by the reviewer. 

Bartlett’s test of sphericity and the Kaiser-Meyer-Olkin

```{r}
library(performance)
check_factorstructure(fa_dat)
```



#### Add 2FA factor scores to df
```{r}
visualization_factor_analysis_2_scores<-psych::factor.scores(fa_dat, visualization_factor_analysis_2)

visualization_factors<-cbind(data_for_fa$Pt_ID,data.frame(visualization_factor_analysis_2_scores$scores))

colnames(visualization_factors)<- c("Pt_ID","Grasp","Appreciate")

cbind(visualization_factors$Pt_ID, full_data$Pt_ID)

full_data<-cbind(full_data,visualization_factors[2:3])
```

## Factor Score Spearman Rank Correlations 
```{r}
fa_scores<-full_data%>%filter(group == "Live")%>%select(Pt_ID,Grasp, Appreciate)

personal_char_list<-df.full%>%select(Pt_ID, age, EC,AIMS, musician_status, fan, familiar_Bach, style_Bach)

fa_scores_and_pers_chars<-full_join(fa_scores, personal_char_list, by = "Pt_ID")

fa_scores_and_pers_chars_nona<-fa_scores_and_pers_chars%>%select(-age, -Pt_ID)%>%na.omit()

names(fa_scores_and_pers_chars_nona)<-c("Grasp \n Factor", "Appreciate \n Factor", "Empathy", "AIMS", "Musician Status", "Fanship", "Familiar Piece", "Familiar Style")
# check number of samples
samples<-nrow(fa_scores_and_pers_chars_nona)
colours_DSQ<-c("#2d769a", "white","#b34036")
title = "Factor Scores"
# title = "Bach Visualization"
subtitle = paste0("Spearman Correlations (BH adj) (n = ", samples, ")")
chart.correlation.spearman(fa_scores_and_pers_chars_nona, colours_DSQ, title, subtitle) #moved the cross-correlation code to the useful_functions script to make this more tidy.

graphname<-paste0("../plots/New_Spearman-Rank-Correlations", title, ".png")
ggsave(graphname, 
       width = 10, 
       height = 9, 
       units = 'cm', 
       dpi = 500)
```

## Assess Group differences: Live versus Online
Compare between live and online complex ratings
IV: group - Live, Online Complex
DV: Grasp, Appreciate

guide: http://www.sthda.com/english/wiki/unpaired-two-samples-wilcoxon-test-in-r

```{r}
complex_data<-full_data%>%filter(type == "Complex" & order != "second")%>%drop_na(c(Grasp, Appreciate))

#check assumptions
complex_data%>%group_by(group)%>%summarise(W = shapiro.test(Grasp)$statistic,
                                          p = shapiro.test(Grasp)$p.value)

complex_data%>%group_by(group)%>%summarise(W = shapiro.test(Appreciate)$statistic,p = shapiro.test(Appreciate)$p.value)

# summarize data
complex_data%>%group_by(group)%>%summarise(n = n())
## Grasp
complex_data%>%group_by(group)%>%summarise(mean = mean(Grasp), SD = sd(Grasp))
## Appreciate
complex_data%>%group_by(group)%>%summarise(mean = mean(Appreciate), SD = sd(Appreciate))

# What are the effects of live versus online on grasping ratings?
wilcox.test(Grasp ~ group, data = complex_data,
                   exact = FALSE)

# What are the effects of live versus online on Appreciate ratings?
wilcox.test(Appreciate ~ group, data = complex_data,
                   exact = FALSE)

```

Visualize
```{r}
plot_data<-complex_data%>%select(Pt_ID,group, Grasp, Appreciate)

long<-plot_data%>%pivot_longer(!c(Pt_ID, group), names_to = "factor", values_to = "response")

ggplot(long, aes(x = group, y = response)) +
  geom_boxplot(fill = "light blue") + 
  geom_jitter(alpha = .5)+
  stat_summary(fun=mean, geom="point", shape=10, size=3.5, color="black") + 
  ggtitle("Live versus Online Complex") + 
  facet_grid(cols = vars(factor))+
  theme_bw() + theme(legend.position="none")
```

## Complexity differences: Online only
Compare between online simple and online complex ratings
IV: type - complex first, Simple first
DV: Grasp, Appreciate

Use only the first video they watched

```{r}
first_data<-full_data%>%filter(group == "Online" & order== "first")%>%drop_na(c(Grasp, Appreciate))

#check assumptions
first_data%>%group_by(type)%>%summarise(W = shapiro.test(Grasp)$statistic,
                                          p = shapiro.test(Grasp)$p.value)

# summarize data
first_data%>%group_by(type)%>%summarise(n = n())
## Grasp
first_data%>%group_by(type)%>%summarise(mean = mean(Grasp), SD = sd(Grasp))
## Appreciate
first_data%>%group_by(group)%>%summarise(mean = mean(Appreciate), SD = sd(Appreciate))

# What are the effects of complex versus simple on Grasping ratings?
wilcox.test(Grasp ~ type, data = first_data,
                   exact = FALSE)

# What are the effects of live versus online on appreciate ratings?
wilcox.test(Appreciate ~ type, data = first_data,
                   exact = FALSE)

```
## Compare the groups: SS, SC, CS, CC
Add subgroup class: SS, SC, CS, CC

DV:Grasp and Appreciate
IV: type -  complex, simple; subgroup
random intercept: Pt_ID
```{r}
## Online: SS, SC, CS, CC
online_dat<-full_data%>%filter(group =="Online")

# join with subgroups df created earlier
online_dat<-online_dat%>%full_join(subgroups, by = "Pt_ID")

online_dat<-online_dat%>%drop_na(Subgroup)
```

Normality is violated in every group
```{r}
#check assumptions
online_dat%>%group_by(Subgroup)%>%summarise(W = shapiro.test(Grasp)$statistic,
                                          p = shapiro.test(Grasp)$p.value)
```

With high sample sizes, the shapiro test is easily violated. Check normality with qqplots.There are minor deviations from normality. Linear mixed effects modeling may be robust against minor deviations to normality, therefore you could still examine if differences could be attributed to subgroup.
```{r}
ggqqplot(online_dat, "Grasp", ggtheme = theme_bw(), facet.by = "Subgroup")
```


### LME

Try with lme4 anyways
#### Grasp
```{r}
baseline<-lmer(Grasp ~ 1 + (1|Pt_ID), data = online_dat, REML = FALSE)
plot(baseline)

type<-lmer(Grasp ~ 1 + type + (1|Pt_ID), data = online_dat, REML = FALSE)
plot(type)

subgroup<-lmer(Grasp ~ 1 + Subgroup + (1|Pt_ID), data = online_dat, REML = FALSE)
plot(subgroup)

anova(baseline, type) # ns
anova(baseline, subgroup) # ns
```
Check qq
```{r}
require(lattice)
qqmath(type, id=0.05)
qqmath(subgroup, id=0.05)
```

#### Appreciate
```{r}
baseline<-lmer(Appreciate ~ 1 + (1|Pt_ID), data = online_dat, REML = FALSE)
plot(baseline)

type<-lmer(Appreciate ~ 1 + type + (1|Pt_ID), data = online_dat, REML = FALSE)
plot(type)

subgroup<-lmer(Appreciate ~ 1 + Subgroup + (1|Pt_ID), data = online_dat, REML = FALSE)
plot(subgroup)

anova(baseline, type) # ns
anova(baseline, subgroup) # ns
```

Visualize
```{r}
plot_data<-online_dat%>%select(Pt_ID,type, Grasp, Appreciate)

long<-plot_data%>%pivot_longer(!c(Pt_ID, type), names_to = "factor", values_to = "response")

ggplot(long, aes(x = type, y = response)) +
  geom_boxplot(fill = "light blue") + 
  geom_jitter(alpha = .5)+
  stat_summary(fun=mean, geom="point", shape=10, size=3.5, color="black") + 
  ggtitle("Online Results") + 
  facet_grid(cols = vars(factor))+
  theme_bw() + theme(legend.position="none")
```

# Did the visualisation support their Grasping?

I need to check if they reported values that are significantly higher than neutral opinion
Scale was 1 to 5 and 3 was in the centre therefore I need to check if the responses on the items were significantly different than 3. 
Bonferroni corrected p-values: multiply by 6.
```{r}
# check assumptions
shapiro.test(fa_scores$Grasp) #violated
shapiro.test(fa_scores$Appreciate) #violated
ggqqplot(fa_scores$Grasp, ylab = "Grasp",
         ggtheme = theme_minimal())
ggqqplot(fa_scores$Appreciate, ylab = "Grasp",
         ggtheme = theme_minimal())

#Therefore you need to conduct the One-Sample Wilcoxon Signed Rank Test in R

#live 
wilcox.test(fa_scores$Grasp, mu = 0, alternative = "two.sided")
wilcox.test(fa_scores$Appreciate, mu = 0, alternative = "two.sided")
```

```{r}
#online first complex
oneSample <-online_dat%>%filter(order == "first")%>%filter(type =="Complex")

# check assumptions
shapiro.test(oneSample$Grasp) #violated
shapiro.test(oneSample$Appreciate) # meets 
ggqqplot(oneSample$Grasp, ylab = "Grasp",
         ggtheme = theme_minimal())
ggqqplot(oneSample$Appreciate, ylab = "Grasp",
         ggtheme = theme_minimal())

wilcox.test(oneSample$Grasp, mu = 0, alternative = "two.sided")
t.test(oneSample$Appreciate, mu = 0, alternative = "two.sided")
```
```{r}
#online first simple
oneSample <-online_dat%>%filter(order == "first")%>%filter(type =="Simple")

# check assumptions
shapiro.test(oneSample$Grasp) #violated
shapiro.test(oneSample$Appreciate) # meets 
ggqqplot(oneSample$Grasp, ylab = "Grasp",
         ggtheme = theme_minimal())
ggqqplot(oneSample$Appreciate, ylab = "Grasp",
         ggtheme = theme_minimal())

wilcox.test(oneSample$Grasp, mu = 0, alternative = "two.sided")
t.test(oneSample$Appreciate, mu = 0, alternative = "two.sided")
```
